  ---
title: "Phenograph clustering"
author: "Xiaofei"
date: "23-05-2024"
output: 
  html_document:
    theme:
      spacelab
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
    number_sections: false
    highlight: textmate 
---

```{r setup, include=FALSE}
#install.packages("remotes")
#remotes::install_github("JinmiaoChenLab/Rphenograph")
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(plyr)
library(stringr)
library(Rphenograph)
```

# Van Maldegem IMC pipeline

The data and methods used in this document were made publicly available by the
authors and can be obtained [here](https://hdl.handle.net/10779/crick.c.5270621.v2).
The purpose of this re-analyis is to become familiar with the pipeline for
analysis of Imaging Mass Cytometry (IMC) data. In this pipeline, the data is
processed in three steps:

- Segmentation (imcyto)
- Normalisation and scaling (R)
- Phenograph clustering (R)

In this exploration we will continue with the third step. During this step,
we perform 'phenotypic' clustering of the celldata with
[Phenograph](https://github.com/dpeerlab/phenograph). Phenograph creates a graph
("network") representing phenotypic similarities between cells, after which it
uses the [Louvain method](https://sites.google.com/site/findcommunities/) to
identify communities in this graph. In this pipeline, we use the [R
implementation of Phenograph](https://github.com/JinmiaoChenLab/Rphenograph)
developed by Jin Miao Chen.


## Set global variables

In this section, we set the input and output directories of this script and
select a data file. This file should contain normalised and scaled IMC data.
Furthermore, we create a vector with the names of the columns that should
be used for clustering.

In this data analysis strategy, we only use lineage markers for clustering to
optimise the clustering results. 



```{r global-variables, include=TRUE, results=FALSE}
# Project folder path:
base = "/mnt/Data1/imcyto/runs/Xiaofei/Megan_IMC"

# Name of input and output directories:
input_path = file.path(base, "Projectfile/Data/Cluster_input")
output_path = file.path(base, "Projectfile/Results/Cluster_output/")

# Name of the file wi th normalised and scaled data:
input_file = "240523_normalised_concatenated_scaled_data_001threshold.csv"

# Load data: 285,296 entries, 49 total columns
celldata = read.csv(file.path(input_path, input_file))

# Names of columns to be used for clustering: 
#COI_names = c("MI_CD103", "MI_CD11b", "MI_CD11c", "MI_CD3e", "MI_CD8a",
 #              "MI_Foxp3", "MI_Ly6G", "MI_MHCII", "MI_Vimentin",'MI_B220',
 #             "MI_PECAM", "MI_aSMA",'MI_CD44')
COI_names = c("MI_CD103", "MI_CD11b", "MI_CD11c", "MI_CD3e", "MI_CD8a",
              "MI_CD45", "MI_CD4", "MI_CD68",  "MI_EpCAM", "MI_F480",
              "MI_Foxp3", "MI_Ly6G", 'MI_CD44',"MI_PVR",
              "MI_PECAM", "MI_aSMA")
#,'MI_CD44',"MI_PVR"
kList = c(50,40,30,25,20) #c(75, 50, 40, 30, 25, 20, 15, 10)
    ```

## Phenograph clustering

### Create object with clustering with multiple k's

```{r}
# Create a clustering object ---------------------------------------------------
# Description : This function creates a dataframe that holds clustering annotation
#               with different k levels
# ---Parameters---
# celldata  --> [dataframe] dataframe where rows denote individual cells, and
#               columns denote mean intensity of each channel
# COInames  --> [vector] names of the columns that should be used for clustering
# Klist     --> [vector] value between 0 and 1 that indicates the percentile
# << Output >>  [dataframe] dataframe where rows denote individual cells, and
#               columns hold cluster annotation with different k levels

create.clustering.object = function(celldata, COInames, Klist) {
  clustering = data.frame("index" = 1:nrow(celldata))
  for (k in Klist) {
    print(Sys.time())
    print(paste0("Current k: ", k))
    # Run Rphenograph on data of select columns
    Rphenograph_out <- Rphenograph(celldata[COI_names], k=k)
    
    # Create a dataframe with the annotation for each cell and sort by index
    annotation = as.vector(membership(Rphenograph_out[[2]]))
    index = as.numeric(names(membership(Rphenograph_out[[2]])))
    clusterDF = data.frame(index, annotation)[order(index),]
    
    # Find the indexes of cells that were not clustered
    allCellIndex = as.numeric(1:nrow(celldata))
    missing = allCellIndex[!allCellIndex %in% clusterDF$index]
    
    # Set the cluster for all unclustered cells to NA:
    if (length(missing) > 0) {
      unassignedCluster = NA
      
      for (n in missing) {
        clusterDF = rbind(clusterDF[1:(n-1),],
                          c(n, unassignedCluster),
                          clusterDF[n:nrow(clusterDF),])
      }
    }
    
    # Add clustering to the celldata dataframe
    if (isTRUE(all.equal(clustering$index, clusterDF$index))){
      colnames = c(colnames(clustering), paste0("kinv", round(1/k * 1000)))
      clustering = cbind(clustering, as.numeric(clusterDF$annotation))
      colnames(clustering) = colnames
    } else {
      stop("Indexes do not correspond")
    }
  }
  
  clustering = subset(clustering, select = -c(index))
  return(clustering)
}

clusteringObject = create.clustering.object(celldata, COI_names, kList)

# Save data with clustering information added
date = str_sub(gsub("-", "", Sys.Date()), -6, -1)
file_path = paste0(output_path, "/", date, "_clustering_k",
                   paste(kList, collapse = "_k"),
                   "_", length(COI_names), "markers.csv")
write.csv(clusteringObject, file = file_path, row.names = F)
print(Sys.time())
```

