---
title: "Preprocessing for Mean"
author: "Xiaofei Yu"
date: "16-05-2024"
output: 
  html_document:
    theme:
      spacelab
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
    number_sections: false
    highlight: textmate 
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(gplots)
library(Rtsne)
library(ggplot2)
library(dplyr)
library(RColorBrewer)
library(stringr)
library(gridExtra)

library(ggridges)
library(ggpubr)
library(scales)
library(clustree)
library(cowplot)
library(umap)
library(lme4)
library(Rphenograph)
```

# IMC Data Analysis

```{r global-variables, include=TRUE, results=FALSE}
# Path for loading data and saving results 
# Project folder path:
base = "/mnt/Data1/groupfebe/runs/Xiaofei/Megan_IMC/"

# Name of input and output directories:
input_path = file.path(base, "Projectfile/Data/Figures_input/")
output_path = file.path(base, "Projectfile/Results/Figures_output/")


# Name of the file wi th normalised and scaled data:
#input_file = "231214_normalised_concatenated_scaled_data_001threshold.csv"

# Load data: 285,296 entries, 49 total columns
#celldata = read.csv(file.path(input_path, input_file))

# Load celldata excluding tSNE coordinates
celldata = read.csv(paste0(input_path, "/240523_normalised_concatenated_scaled_data_001threshold.csv"))

# Add clustering and t-SNE columns to datafiles
tSNECSV = read.csv(paste0(input_path, "/240523_tSNE.csv"))
clustCSV = read.csv(paste0(input_path,"/240523_clustering_k50_k40_k30_k25_k20_16markers.csv"))
#clustCSV = clustCSV$clusterk25
celldata = cbind(celldata,tSNECSV, clustCSV)
MacDendCD = read.csv(paste0(input_path,"MacDendCD.csv"))

imagemetadata = read.csv(file.path(input_path, "/imagemetadata.csv"))
#change celldata experimental group name for making plot 
celldata$ExpGroup[celldata$ExpGroup=="CONTROL"]<-"MRTX+PD1"
celldata$ExpGroup[celldata$ExpGroup=="TREATMENT"]<-"MRTX+PD1+CTLA-4"
# Create a vector with the original ROI_IDs
# Replace all instances of CONTROL with MRTX+PD1 in the ROI_ID column
celldata$ROI_ID <- gsub("CONTROL", "MRTX+PD1", celldata$ROI_ID)
celldata$ROI_ID <- gsub("TREATMENT", "MRTX+PD1+CTLA-4", celldata$ROI_ID)
celldata$filename <- gsub("CONTROL", "MRTX+PD1", celldata$filename)
celldata$filename <- gsub("TREATMENT", "MRTX+PD1+CTLA-4", celldata$filename)


# celldata include anncluster,tSNE,annotation information
#write.csv(celldata,file = paste0(output_path,"/240528celldata.csv"))
# celldata = read.csv(paste0(input_path, "/240528celldata.csv"))
Umap_myeloidCD = read.csv(paste0(input_path, "/Umap_myeloidCD.csv"))
```


# Define function  
## 1  get cluster center,2  make matrix, 3 make heatplot 
##  define plot. heatmap ; plot.scatter ; plot. boxplot function
## If want to change something on plot  ,then copy from here, or just change here, 

```{r}
# Plot tSNE coloured by Phenograph cluster2 & clustername4 
get.cluster.centers = function(celldata, tSNEaxes, clustering){
  celldata = celldata[!is.na(celldata[,clustering]),]
  clusterCenters = data.frame()
  for (cl in unique(celldata[,clustering])){
    x = median(celldata[,tSNEaxes[1]][which(celldata[,clustering] == cl)])
    y = median(celldata[,tSNEaxes[2]][which(celldata[,clustering] == cl)])
    clusterCenters[cl, "cl"] = cl
    clusterCenters[cl, "x"] = x
    clusterCenters[cl, "y"] = y
  }
  return(clusterCenters)
}

make.heatmap.matrix = function(celldata, markerlist, cluster){
  heatmap_df = data.frame()
  
  # Leave out unclustered cells (cluster == NA)
  celldata = celldata[!is.na(celldata[,cluster]),]
  
  for (cl in unique(celldata[,cluster])){
    md = celldata[celldata[,cluster] == cl,]
    for (marker in markerlist){
      heatmap_df[cl, marker] =  mean(md[,paste("MI_", marker, sep = "")])
      heatmap_df[cl, "clustername"] = cl
    }
  }
  
  row.names(heatmap_df) = heatmap_df$clustername
  heatmap_df = t(as.matrix(subset(heatmap_df, select = -clustername)))
  
  return(heatmap_df)
}



plot.heatmap2 = function(celldata, markerlist, cluster, scale="column", title,
                         margins=c(4,5), srtCol=0, Rowv=NULL, Colv=NULL,
                         barplot = c(), type = "stack", transpose = F
){
  matrix = make.heatmap.matrix(celldata, markerlist, cluster)
  
  if (transpose) {matrix = t(matrix)}
  
  hm = heatmap.2(matrix,
                 dendrogram = "none", Rowv = Rowv, Colv = Colv, 
                 scale = scale, trace = "none", 
                 col = colorRampPalette(brewer.pal(8, "Blues"))(25),
                # col = colorRampPalette(c("blue","white","red"))(25),
                 
                 main = title, margins = margins,
                 cexRow = 1, cexCol = 1, srtCol = srtCol,
                 keysize = .8, density.info = "none", key.par=list(mar = c(4,0.5,2,0.7)),
                 lmat = matrix(data = c(0,0,0,4,1,1,3,1,1,2,2,2), ncol = 4),
                 lwid = c(0.5, 1.2, 6.5, 0.01), lhei = c(2, 1.8, 5))
  
  if (length(barplot) != 0) {
    for (i in 1:length(barplot)) {
      df = as.data.frame(table(paste(celldata[!is.na(celldata[,cluster]), cluster],
                                     celldata[,barplot[i]], sep = "x_x")))
      df = cbind(str_split_fixed(df[,1], "x_x", n = 2), df[,2])
      df = data.frame(as.numeric(df[,1]), df[,2], as.numeric(df[,3]))
      colnames(df) = c("cluster", "ROI_ID", "Cell_count")
      df$cluster = factor(df$cluster, levels = hm$colInd)
      
      p = ggplot(df, aes(fill=ROI_ID, y=Cell_count, x=cluster)) + 
        geom_bar(position=type, stat="identity", orientation = "x") +
        ggtitle(paste0("Cell count by ", barplot[i])) + theme_classic() +
        theme(plot.title = element_text(hjust = 0.5, size = 12, face = "bold"),
              axis.text.x = element_text(angle = 90, vjust = .5),
              plot.margin = unit(c(0.02, 0, .03, 0), units = "npc"))
      print(p)
    }
  }
}

plot.scatter = function(celldata, cluster, clusterN, Xmarker, Ymarker,
                        main = "", dim = 2, layout = "ghostvar") {
  # Select celldata of the selected clusters
  celldata = celldata[celldata[,cluster] %in% clusterN,]
  
  # Convert the markers into symbols to be used by ggplot
  x = sym(paste0("MI_", Xmarker))
  y = sym(paste0("MI_", Ymarker))
  
  # Set the nbin based on the celldata size
  nbin = ifelse(nrow(celldata)>200000, 1024, 
                ifelse(nrow(celldata)>20000, 512, 256))
  
  # Color by density
  colramp = colorRampPalette(rev(brewer.pal(n = 9, name = "YlGn")[3:9]))
  color = densCols(x = celldata[,paste0("MI_", Xmarker)],
                   y = celldata[,paste0("MI_", Ymarker)],
                   colramp = colramp, nbin = nbin)
  
  ggplot(data = celldata, aes(x = !!x, y = !!y)) +
    geom_point(shape = I(20), size = 1.5, aes(x = !!x, y = !!y, col = color)) +
    scale_x_continuous(limits = c(0,dim), expand = c(0,0),
                       oob = scales::squish) +
    scale_y_continuous(limits = c(0,dim), expand = c(0,0),
                       oob = scales::squish) + scale_color_identity() + 
    ggtitle(main) + ylab(Ymarker) + xlab(Xmarker) + theme_classic() + 
    theme(plot.title = element_text(hjust = 0.5, size = 12, face = "bold"),
          plot.margin = unit(c(0.05, 0, .05, 0), units = "npc"),
          axis.title = element_text(size = 8, face = "bold"),
          axis.text = element_text(size = 8, colour = "black"),
          panel.background = element_rect(fill = "grey95"), aspect.ratio = 1)
}


plot.scatter3 = function(celldata, cluster, clusterN, Xmarker, Ymarker,
                         colorMarker, main = "", dim = 2, layout = "small") {
  # Select celldata of the selected clusters
  celldata = celldata[celldata[,cluster] %in% clusterN,]
  
  # Convert the markers into symbols to be used by ggplot
  x = sym(paste0("MI_", Xmarker))
  y = sym(paste0("MI_", Ymarker))
  
  if(grepl("ROI", colorMarker, fixed = T)) {
    c = sym(colorMarker)
  } else {
    c = sym(paste0("MI_", colorMarker))
  }
  
  # Change the size of the legend based on the layout parameter
  if (layout == "small") {
    keySize = unit(.013, "npc")
    keyPos = c(0.71, 0.90)
    keyMar = margin(t = 0.005,r = 0.005,b = 0.005,l = 0.005, unit = "npc")
    keyTextSize = 4
  } else {
    keySize = unit(.03, "npc")
    keyPos = c(0.75, 0.92)
    keyMar = margin(t = 0.01, r = 0.01, b = 0.01, l = 0.01, "npc")
    keyTextSize = 6
  }
  
  ggplot(data = celldata, aes(x = !!x, y = !!y)) +
    geom_point(shape = I(20), size = 1.5, aes(x = !!x, y = !!y, color = !!c)) +
    scale_color_gradientn(colours = rainbow(5)) + labs(color = colorMarker) +
    scale_x_continuous(limits = c(0,dim), expand = c(0,0),
                       oob = scales::squish) +
    scale_y_continuous(limits = c(0,dim), expand = c(0,0),
                       oob = scales::squish) +
    ggtitle(main) + xlab(Xmarker) + ylab(Ymarker) + theme_classic() +
    
    theme(plot.title = element_text(hjust = 0.5, size = 12, face = "bold"),
          plot.margin = unit(c(0.02, 0, .03, 0), units = "npc"),
          axis.title = element_text(size = 8, face = "bold"),
          axis.text = element_text(size = 8, colour = "black"),
          legend.title = element_text(size = 8, face= "bold"),
          legend.key.size = keySize, legend.position = keyPos,
          legend.margin = keyMar, legend.text = element_text(size = keyTextSize),
          legend.background = element_rect(fill = "grey80"),
          legend.justification = "center", legend.direction = "horizontal",
          panel.background = element_rect(fill = "grey95"), aspect.ratio = 1)
}



plot.boxplot = function(celldata, factor, comparisons, title) {
  factor = sym(factor)
  ggplot(celldata, aes(x = ExpGroup, y = !!factor, colour = ExpGroup)) +
    
    geom_boxplot() + geom_dotplot(aes(fill = MouseID),method = "histodot", binaxis='y',
                                  stackdir='center', dotsize = 0.8) +
    scale_y_continuous(labels = percent) +
    scale_x_discrete(limits=c("CONTROL", "TREATMENT",'ETP')) +
    stat_compare_means(comparisons = comparisons, label = "p.signif") +
    
    theme_classic() + ggtitle(title) +
    theme(plot.title = element_text(hjust = 0.5, size = 20, face = "bold"),
          plot.background = element_blank(),
          axis.title = element_blank(),
          axis.text = element_text(size = 14, face = "bold"),
          legend.position = "none",
          legend.title = element_blank(),
          legend.text = element_text(size = 15))
  
  ggsave(filename = paste0(date, "Boxplot_", factor, ".jpeg"),
         path = file.path(output_path, "Boxplots"), bg = "white",
         width = 2100, height = 1700, units = "px")
  
}


```
####filter celldata without ETP group
```{r}
celldata<-celldata[(celldata$ExpGroup %in% c("MRTX+PD1","MRTX+PD1+CTLA-4")),]
unique(celldata$ExpGroup)
date <- Sys.Date()
output_file = paste0(output_path, "/",date, "_celldata.csv")
output_file
write.csv(celldata, file = output_file, row.names = F) 
input_file = paste0(input_path, date, "_celldata.csv")
input_file
write.csv(celldata, file = input_file, row.names = F) 
```



## Plot new t-SNE   No need to change
# t-SNE
##define plot tSNE function
```{r plot tSNE function}
mycolors = sample(colorRampPalette(brewer.pal(8, "Set1"))(30))
# Plot the generated t-SNE with the current data
plot.tSNE = function(celldata, tSNEaxes, clustering, lvls, title,
                     labelClusters = T, savetSNE = F) {
  
  if (labelClusters) {
    clusterCenters = get.cluster.centers(celldata, tSNEaxes, clustering)
  }
  
  tSNEx = sym(tSNEaxes[1])
  tSNEy = sym(tSNEaxes[2])
  clustering = sym(clustering)
  
  if (missing(lvls)) {
    lvls = clustering
  }
  
  p = ggplot(celldata[sample(nrow(celldata), 20000),]) +
    geom_point(size = 0.1,
               aes(x = !!tSNEx, y= !!tSNEy,
                   colour = factor(!!clustering,
                                   levels = sort(unique(!!lvls))))) +
    scale_color_manual(values = mycolors) +
    
    # Plot the cluster centers
    {if (labelClusters) {
      list(geom_point(data = clusterCenters, size = 6, aes(x=x, y=y),
                      colour = "black", pch = 21, fill = "white",
                      show.legend = F),
           geom_text(data = clusterCenters, stat = "identity", show.legend = F,
                     mapping = aes(x = x, y = y, label = cl, size = 12),
                     colour = "black", size = 3))}} +
    
    theme_classic() +
    theme(plot.title = element_text(size = 20, hjust = 0.5),
          legend.title = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          legend.text = element_text(size = 10)) +
    guides(colour = guide_legend(override.aes = list(size=5))) +
    ggtitle(title) 
  
  if (savetSNE) {
    date = str_sub(gsub("-", "", Sys.Date()), -6, -1)
    title_name = gsub(pattern = " ", replacement =  "_", x = title)
    filename = paste0(date, "_t-SNE_", title_name, ".jpeg")
    ggsave(plot = p, device = "jpeg", bg = "white",
           width = 2500, height = 1700, dpi = 300, units = "px",
           path = output_path, filename = filename)
  }
  
  return(p)
}
```
##Visulaize TSNE  (Check mixture)
```{r}
plot.tSNE(celldata, c("tSNE1", "tSNE2"), "ExpGroup",
          title = "Experimental model", labelClusters = F, savetSNE = F)
plot.tSNE(celldata, c("tSNE1", "tSNE2"), "ROI_ID", savetSNE = F,
          title = "ROI ID", labelClusters = F)
plot.tSNE(celldata, c("tSNE1", "tSNE2"), "kinv33", labelClusters = F,
          unique(celldata$kinv33), title = "Annotated cell types",
          savetSNE = T)


# 240523 change code due to the naming of models,,By using this approach, you ensure precise matching of the  names, capturing only the exact  values that match "MRTX+PD1" and "MRTX+PD1+CTLA-4" without overlapping.modelROI_ID  regexpr 返回正则表达式模式在 text 向量的每个元素中的第一个匹配的起始位置。如果没有找到匹配，则返回 -1
plottingtSNE = function(model) {
  levels = unique(celldata$ROI_ID)[regexpr(pattern = paste0(model,"$"), text = unique(celldata$ROI_ID))>0]
  # levels = grep(pattern = paste0(model), x = unique(celldata$ROI_ID), value = T,fixed = TRUE)
  p = plot.tSNE(celldata, c("tSNE1", "tSNE2"), "ROI_ID", lvls = levels,
                title = paste0(model, "_ROIs"), labelClusters = F, savetSNE = T)
  return(list(p))
}

exp_models = c("PD1","CTLA-4")
# unique(celldata$ROI_ID)[regexpr(pattern = paste0(model,"$"), text = unique(celldata$ROI_ID))>0]

sapply(X = exp_models, FUN = plottingtSNE)

```

# Heatmap

**Figure 3b**\
# Generate a heatmap with markers versus clusters.

```{r create heatmap with markers}
# lineage markers only (16 nkp46-B220-Vimentin+CD8-):
#B220 binding to nuclei ,so maybe manually choose cluster for B220
lin_markers = c("CD45", "CD3e", "CD4", "Foxp3",'CD8a',"MHCII",
                "CD11c","CD103", "F480", "CD68", "CD11b", "Ly6G", "PECAM",
                "EpCAM", "aSMA","B220","PVR","CD44")
# dexlude DNA ,domain, BCK, 
excl_cols = c( "MI_Xenon131", "MI_Xenon132", "MI_DNA1",
              "MI_DNA2", "MI_Argon80")

all_markers = grep("MI_", names(celldata), value = T)
all_markers = all_markers[!all_markers %in% excl_cols] %>%
  str_remove(pattern = "MI_")

plot.heatmap2(celldata, lin_markers, "kinv33", title = "Cluster k30 lineage markers",
              Colv = T,  Rowv = T, srtCol = 90, barplot = c("ExpGroup", "ROI_ID"))
plot.heatmap2(celldata, all_markers, "kinv33", title = "Cluster k30 all markers",
              Colv = T,  Rowv = T, srtCol = 90, barplot = c("ExpGroup", "ROI_ID"))
plot.heatmap2(celldata, all_markers, "kinv33", title = "ROI_allmarkers",margins = c(6,5),
              Colv = T, Rowv = T,srtCol = 90,scale="row")
plot.heatmap2(celldata, all_markers, "kinv33", title = "ROI_allmarkers",margins = c(6,5),
              Colv = T, Rowv = T,srtCol = 90)


plot.heatmap2(celldata, lin_markers, "kinv33", title = "Cluster k30",
              Colv = T,Rowv = T, srtCol = 90)
plot.heatmap2(celldata, all_markers, "kinv33", title = "Cluster k30",
              Colv = T, srtCol = 90)

plot.heatmap2(celldata, lin_markers, "ROI_ID", title = "ROI_linmarkers",margins = c(10,5),
              Colv = T, srtCol = 90)

plot.heatmap2(celldata, lin_markers, "ROI_ID", title = "ROI_linmarkers",margins = c(10,5),
              Colv = T, srtCol = 90,scale="row")


plot.bp = function(clustern) {
  filename = file.path(file.path(output_path, "Figures/Cluster"),
                       paste0(date, "_cluster", clustern, ".jpeg"))
  jpeg(filename = filename, width = 700, height = 600, units = "px")
  par(mar=c(11,4,1,1))
  barplot(table(celldata[celldata$kinv33 == clustern, "filename"]),
          las = 2, cex.names = 0.8)
  dev.off()
}

date = str_sub(gsub("-", "", Sys.Date()), -6, -1)
sapply(c(1:25), plot.bp)

#TODO bargraph divide by amount of cells per group (normalise)
#TODO autosave heatmaps + bargraphs


#plot.heatmap2(celldata, lin_markers, "kinv33", title = "Cluster k30",
#Colv = T, srtCol = 90, barplot = c("ExpGroup", "ROI_ID"))
#plot.heatmap2(celldata, lin_markers, "kinv33", title = "Cluster k30",
#Colv = T, srtCol = 90, barplot = c("ExpGroup", "ROI_ID"), scale = "row")

```

# Check the cluster threshold for annotation
```{r check threshold}
celldata$kin
plot.scatter(celldata, "kinv40", 4, "CD11c", "F480",dim = 1)
plot.scatter(celldata, "kinv40", 8, Xmarker = "Foxp3", Ymarker = "CD4")
plot.scatter(celldata, "kinv40", 12, Xmarker = "CD4", Ymarker = "CD8a",dim=1)

plot.scatter(celldata, "anncluster", "07A_Tregss", Xmarker = "Foxp3", Ymarker = "EpCAM")
plot.scatter(celldata, "anncluster", "02A_Tumour", Xmarker = "EpCAM", Ymarker = "CK19")
plot.scatter(celldata, "anncluster", "19A_Tumour", Xmarker = "EpCAM", Ymarker = "CK19")
plot.scatter(celldata, "annotation", "Tumour", "EpCAM", "CK19",
             main = "Tumour", dim = 2.5)
plot.scatter(celldata, "annotation", "T cells CD4", "CD3", "Foxp3",
             main = "T Cells CD4", dim = 5)
# T cells
plot.scatter(celldata, "annotation", c("T cells CD4", "Tregs" ), "CD4", "Foxp3",
             main = "T Cells", dim = 4)
plot.scatter(celldata, "annotation", "T cells CD4", "CD3e", "Foxp3",
             main = "T Cells CD4", dim = 5)
plot.ridge(celldata, "CD4", "annotation", c("T cells CD4", "T cells CD8a", "Macrophages"))

plot.ridge(celldata, "CD3e", "annotation", unique(celldata$annotation))

# Fibroblasts?
plot.scatter(celldata, "kinv100", 7, "aSMA", "Thy1_2", main = "Cluster 7")

# Epithelium / macrophage?
plot.scatter(celldata, "kinv100", 27, "MHCcII", "CD11c", main = "Cluster 27")
plot.scatter(celldata, "kinv100", 27, "F480", "CD68", main = "Cluster 27")
plot.scatter3(celldata, "kinv100", 27, "aSMA", "PancK", "F480",
              main = "Cluster 27", layout = "big")

plot.scatter(celldata, "kinv100", 27, "aSMA", "PancK", #"ROI_ID",
             main = "Cluster 27", layout = "big")

# Fix margins for combined figure:
p1 = plot.scatter3(celldata, "kinv40", 26, "CD4", "Foxp3", "ExpGroup")
p2 = plot.scatter3(celldata, "kinv50", 24, "F480", "aSMA", "EPCAM")
p3 = plot.scatter3(celldata, "kinv50", 24, "F480", "CD68", "PECAM")
p4 = plot.scatter(celldata, "kinv50", 24, "EPCAM", "F480")
plot1 = ggarrange(p1, p2, p3, p4, ncol = 3, nrow = 2)
annotate_figure(plot1, top = text_grob("Cluster24", face = "bold", size = 14))

```




##manual annotation
```{r manual annotation}
# Manually assign cluster labels (current: k30)
annotate.cluster = function(clusterCol) {
  for (cluster in unique(clusterCol)) {
    newname = switch(cluster,
                     "01" = "01_T cells CD8",
                     "02" = "02_Endothelium", 
                     "03A" = "03A_Macrophages type 1",
                     "03B" = "03B_Macrophages other",
                     "04" = "04_Macrophages type 2",
                     "05" = "05_Fibroblasts",
                     "06" = "06_Tumour",
                     "07" = "07_Dendritic cells CD103",
                     "08" = "08_Tumour",
                     "09" = "09_Endothelium",
                     "10" = "10_Tumour", 
                     "11" = "11_Tumour",
                     "12B" = "12B_T cells CD4",
                      "12A" = "12A_T reg cells",
                     "13" = "13_Tumour",
                     "14" = "14_Neutrophils",
                     "15A" = "15A_Endothelium", 
                     "15B" = "15B_Epithelium", 
                     "16" = "16_Endothelium",
                     "17" = "17_Endothelium",
                     "18A" = "18A_Macrophages other",
                     "18B" = "18B_Tumour",
                     "19" = "19_Tumour",
                     "20" = "20_Endothelium",
                     "21" = "21_Epithelium",
                     "22" = "22_Epithelium",
                     "23A" = "23A_Dendritic cells",
                     "23B" = "23B_Macrophages type 2",
                     "24" = "24_T cells CD4",
                     "25" = "25_Tumour",
                     "26" = "26_B cells",
                     "00_Undefined"
    )
    clusterCol[clusterCol == cluster] = newname
  }
  return(clusterCol)
}

# Generate a column "anncluster" that makes a string out of the cluster numbers
celldata$anncluster = str_pad(celldata$kinv33, 2, side="left", pad="0")

# # Clustering annotation
# 
# In this section, the clusters as defined in the previous step are manually
# annotated. First, clusters that consist of two different cell types
# are split based on (a combination of) marker values.
# 
# Then, each (sub)cluster is mannually annotated based on the expression data (?).
# The following columns are added to the datafile:
# 
# - anncluster: holds the (sub)cluster identifier + _ + annotation;
# - annotation: holds the cell type annotation.



# # Split cluster 17 into Epithelium and Endothelium 
celldata[celldata$anncluster == "15"
         & celldata$MI_PECAM >= 0.04 , #& celldata, $MI_EpCAM  < 0.2,
         "anncluster"] = "15A"
celldata[celldata$anncluster == "15", "anncluster"] = "15B"
# # pick out CD8a from cluster 10?
#split cluster 18
celldata[celldata$anncluster == "18"
         & celldata$MI_F480 > 0.5, #& celldata, $MI_EpCAM  < 0.2, 
         "anncluster"] = "18A"
celldata[celldata$anncluster == "18", "anncluster"] = "18B"
#split 13
celldata[celldata$anncluster == "12"
         & celldata$MI_Foxp3>0.01, 
         "anncluster"] = "12A"
celldata[celldata$anncluster == "12", "anncluster"] = "12B"
# split cluster 23 
celldata[celldata$anncluster == "23"
 & celldata$MI_CD11c>0.2, 
    "anncluster"] = "23A"
celldata[celldata$anncluster == "23", "anncluster"] = "23B"
# split macrophages 
celldata[celldata$anncluster == "03"
         & celldata$MI_CD206<0.4, 
         "anncluster"] = "03A"
celldata[celldata$anncluster == "03", "anncluster"] = "03B"
# choose B cell cluster from all 
celldata[celldata$MI_B220 > 0.7, "anncluster"] = "26"   



# Generate a column "cluster2" that makes a string out of the cluster numbers
celldata$anncluster = annotate.cluster(celldata$anncluster)

celldata$annotation = str_split(celldata$anncluster,
                                pattern = "_", simplify = TRUE)[,2]
```



# recluster for myleoid annotation
```{r}
#   extract subgroup myeloidCD  from celldata
#Myeloidclusters = c(23,03,07,04,18)
Myeloidclusters = c(grep(pattern = "Macrophage", unique(celldata$anncluster), value = T),
                grep(pattern = "Dendritic", unique(celldata$anncluster), value = T))
myeloidCD <- celldata[celldata$anncluster %in% Myeloidclusters, ]
# write.csv(x = myeloidCD, file = file.path(output_path, "myeloidCD.csv"), row.names = T)
#  recluster myeloids
COI_names = c("MI_MHCII", "MI_F480", "MI_CD103", "MI_PDL1", "MI_CD86",
                       "MI_CD68", "MI_CD206", "MI_pS6", "MI_CD11c", "MI_CD11b","MI_CD44")
#kList = c(110,130,150,170)
#kList = c(200,250,300)
kList = c(80,100,120,140,160,400,500)
create.clustering.object = function(myeloidCD, COInames, Klist) {
  clustering = data.frame("index" = 1:nrow(myeloidCD))
  for (k in Klist) {
    print(Sys.time())
    print(paste0("Current k: ", k))
    # Run Rphenograph on data of select columns
    Rphenograph_out <- Rphenograph(myeloidCD[COI_names], k=k)
    # Create a dataframe with the annotation for each cell and sort by index
    annotation = as.vector(membership(Rphenograph_out[[2]]))
    index = as.numeric(names(membership(Rphenograph_out[[2]])))
    clusterDF = data.frame(index, annotation)[order(index),]
    
    # Find the indexes of cells that were not clustered
    allCellIndex = as.numeric(1:nrow(myeloidCD))
    missing = allCellIndex[!allCellIndex %in% clusterDF$index]
    
    # Set the cluster for all unclustered cells to NA:
    if (length(missing) > 0) {
      unassignedCluster = NA
      
      for (n in missing) {
        clusterDF = rbind(clusterDF[1:(n-1),],
                          c(n, unassignedCluster),
                          clusterDF[n:nrow(clusterDF),])
      }
    }
    
    # Add clustering to the celldata dataframe
    if (isTRUE(all.equal(clustering$index, clusterDF$index))){
      colnames = c(colnames(clustering), paste0("kinv", round(1/k * 1000)))
      clustering = cbind(clustering, as.numeric(clusterDF$annotation))
      colnames(clustering) = colnames
    } else {
      stop("Indexes do not correspond")
    }
  }
  
  clustering = subset(clustering, select = -c(index))
  return(clustering)
}

clusteringObject = create.clustering.object(myeloidCD, COI_names, kList)

# Save data with clustering information added
date = str_sub(gsub("-", "", Sys.Date()), -6, -1)
file_path = paste0(output_path, "/", date, "_clustering_k",
                   paste(kList, collapse = "_k"),
                   "_", length(COI_names), "markers.csv")
write.csv(clusteringObject, file = file_path, row.names = F)
print(Sys.time())

#
myeloid_clustCSV = read.csv(paste0(input_path,"/240528_clustering_k80_k100_k120_k140_k160_k400_k500_11markers.csv"))
myeloidCD = cbind(myeloidCD, myeloid_clustCSV)
# remove useless kinv
myeloidCD <- myeloidCD[, !names(myeloidCD) %in% c("kinv2","kinv2.1","kinv7","kinv10","kinv12","kinv3","kinv4","kinv20","kinv25","kinv40","kinv50","kinv5.1","kinv5.2","kinv6","kinv8")]


lin_myeloid_markers = c("MHCII", "F480", "CD103", "PDL1", "CD86",
                       "CD68", "CD206", "pS6", "CD11c", "CD11b","CD44")


plot.heatmap2(myeloidCD, lin_myeloid_markers, "kinv5", title = "Cluster k200",
              Colv = T, srtCol = 90,scale = "row")


### Next  I choose k200  for heatmap
###2406026 Xiaofei reorder the code 
# Manually assign cluster labels (current: k200,kinv5)
# Generate a column "anncluster" that makes a string out of the cluster numbers
myeloidCD$anncluster = str_pad(myeloidCD$kinv5, 2, side="left", pad="0")
print(unique(myeloidCD$anncluster))
myeloidCD[myeloidCD$anncluster == "04"
         & (myeloidCD$MI_MHCII > 0.9|myeloidCD$MI_CD11c>1) , #& celldata, $MI_EpCAM  < 0.2,
         "anncluster"] = "04A"
myeloidCD[myeloidCD$anncluster == "04"& (myeloidCD$MI_F480 > 0.5), "anncluster"] = "04B"
myeloidCD[myeloidCD$anncluster == "04"& (myeloidCD$MI_F480 <= 0.5), "anncluster"] = "04C"
print(unique(myeloidCD$anncluster))
annotate.cluster = function(clusterCol) {
  for (cluster in unique(clusterCol)) {
    newname = switch(cluster,
                     "01" = "01_Macrophages type 1",
                     "02" = "02_Macrophages type 2",
                     "03" = "03_Macrophages type 2",
                     "04A" = "04A_Dendritic cells",
                     "04B" = "04B_Macrophages type 2",
                     "04C" = "04C_Macrophages type 1",
                     "05" = "05_Macrophages type 2",
                     "06" = "06_Macrophages type 1",
                     "07" = "07_Dendritic cells CD103",
                     "08" = "08_Macrophages type 2",
                     "09" = "09_Macrophages type 2",
                     "10" = "10_Macrophages type 1",
                     "00_Undefined"
    )
    clusterCol[clusterCol == cluster] = newname
  }
  return(clusterCol)
}
# Apply cluster annotation
myeloidCD$anncluster = annotate.cluster(myeloidCD$anncluster)
print(unique(myeloidCD$anncluster))
# Extract annotation from the updated cluster column
myeloidCD$annotation = str_split(myeloidCD$anncluster,
                             pattern = "_", simplify = TRUE)[,2]
print(unique(myeloidCD$annotation))
# annotated macrophages

plot.heatmap2(myeloidCD, lin_myeloid_markers, "anncluster", title = "Annotated clusters",
              margins = c(10, 5), srtCol = 60, Colv = T)
plot.heatmap2(myeloidCD, lin_myeloid_markers, "annotation",
              title = "Cell type expression pattern",
              margins = c(10, 9), srtCol = 55)

## make a new umap for new myeloidCD
  MyeloidClusters = unique(myeloidCD$anncluster)
  MyeloidColumns = c("MI_MHCII", "MI_F480", "MI_CD103", "MI_PDL1", "MI_CD86",
                         "MI_CD68", "MI_CD206", "MI_pS6", "MI_CD11c", "MI_CD11b",'MI_Sirpa',"MI_CD44") #, "MI_LY6G",  "MI_CD44")
  Umap_myeloidCD_newDC = generate.UMAP(datatable = myeloidCD, clusters = MyeloidClusters,
                            markers = MyeloidColumns,
                            filename = "Umap_myeloidCD_newDC.csv")

ggplot(Umap_myeloidCD_newDC,aes(x=umap1, y=umap2, colour = anncluster))+geom_point()
ggplot(Umap_myeloidCD_newDC,aes(x=umap1, y=umap2, colour = annotation))+geom_point()

#febe map maker on generted umap figures. 
ggplot(Umap_myeloidCD,aes(x=umap1, y=umap2, colour = anncluster)) + 
    geom_point() + 
    geom_point(data = Umap_myeloidCD[which(Umap_myeloidCD$anncluster == "04_Dendritic cells"& Umap_myeloidCD$MI_MHCII>0.9),], colour = "black")


##save new myeloidCD

date <- Sys.Date()
output_file = paste0(output_path, date, "_myeloidCD.csv")
output_file
write.csv(myeloidCD, file = output_file, row.names = F) 

```
#Xiaofei:Get new annotation because of the recluster of macrophages. 
```{r}

#ifelse(new_anncluster,anncluster),match -->index,anncluster[index]--> needed names of anncluster,cell_ID %in% Umap_myeloidCD$cell_ID-->which row is True
print(unique(celldata$anncluster))
celldata<-celldata %>% 
  dplyr::mutate(anncluster=ifelse(cell_ID %in% myeloidCD$cell_ID,
                      myeloidCD$anncluster[match(cell_ID,myeloidCD$cell_ID)],
                      anncluster
                      ))
print(unique(celldata$anncluster))
## To keep anncluster number unique 
celldata[which(celldata$anncluster == "01_Macrophages type 1"), "anncluster"] = "31_Macrophages type 1"
celldata[which(celldata$anncluster == "02_Macrophages type 2"), "anncluster"] = "32_Macrophages type 2"
celldata[which(celldata$anncluster == "03_Macrophages type 2"), "anncluster"] = "33_Macrophages type 2"
celldata[which(celldata$anncluster == "05_Macrophages type 2"), "anncluster"] = "35_Macrophages type 2"
celldata[which(celldata$anncluster == "06_Macrophages type 1"), "anncluster"] = "36_Macrophages type 1"
celldata[which(celldata$anncluster == "08_Macrophages type 2"), "anncluster"] = "38_Macrophages type 2"
celldata[which(celldata$anncluster == "09_Macrophages type 2"), "anncluster"] = "39_Macrophages type 2"
celldata[which(celldata$anncluster == "10_Macrophages type 1"), "anncluster"] = "40_Macrophages type 1"
print(unique(celldata$anncluster))

plot.heatmap2(celldata, all_markers, "anncluster", title = "Annotated clusters",
              margins = c(10, 5), srtCol = 60, Colv = T)
plot.heatmap2(celldata, lin_markers, "anncluster", title = "Annotated clusters",
              margins = c(10, 5), srtCol = 60, Colv = T)
plot.heatmap2(celldata, lin_markers, "annotation",
              title = "Cell type expression pattern",
              margins = c(10, 9), srtCol = 55)

ggplot(Umap_myeloidCD_newDC,aes(x=umap1, y=umap2, colour = anncluster))+geom_point()
ggplot(Umap_myeloidCD_newDC,aes(x=umap1, y=umap2, colour = annotation))+geom_point()
ggplot(Umap_myeloidCD_newDC,aes(x=umap1, y=umap2, colour = ExpGroup))+geom_point()

#####240626
MacDendCellColumns = c("MI_MHCII", "MI_F480", "MI_CD103", "MI_PDL1", "MI_CD86",
                       "MI_CD68", "MI_CD206", "MI_pS6", "MI_CD11c", "MI_CD11b","MI_CD44")
for (i in MacDendCellColumns){
  marker = str_remove(i, "MI_")
  p = ggplot(Umap_myeloidCD_newDC,
             aes(x=umap1, y=umap2, colour = get(i))) + 
    geom_point(size = 0.3) + ggtitle(paste0("Myeloid ", marker, " expression")) + 
    theme_classic() + labs(x = "UMAP 1", y = "UMAP 2") +
    theme(plot.background = element_blank(),
          axis.title = element_text(size = 15),
          legend.text = element_text(size = 15),
          legend.title = element_blank(),
          plot.title = element_text(hjust = 0.5, size = 25, face = "bold"),
          panel.grid = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank()) +
    scale_color_distiller(palette = "Spectral", limits = c(0,3))
  ggsave(filename = paste0("UMAP_Myeloid_", marker, "_expression.jpeg"),
         path = file.path(output_path, "MyeloidUMAP"), bg = "white",
         width = 2100, height = 1700, units = "px")
  print(p)
}

celldata = read.csv(paste0(input_path, "2024-06-26_celldata.csv"))
celldata$annotation = str_split(celldata$anncluster,
                                pattern = "_", simplify = TRUE)[,2]


date <- "2024_06_26"
output_file = paste0(output_path, date, "_celldata.csv")
output_file
input_file = paste0(input_path, date, "_celldata.csv")
input_file
write.csv(celldata, file = input_file, row.names = F) 


```






```


```{r new UMAPs}
get.clusters = function(patterns) {
  clusters = sapply(X = patterns, FUN = grep, x = unique(celldata$anncluster), value = T)
  clusters = unlist(clusters, use.names = F)
  return(clusters)
}
if (F) {
  # T cells
  TCellClusters = get.clusters(patterns = c("T cell",'T reg'))
  TCellMarkers = c("MI_CD103", "MI_CD3e", "MI_CD45", "MI_CD4", "MI_CD8a", "MI_Foxp3",
                   "MI_Ki67", "MI_LAG3", "MI_PD1","MI_Casp3","MI_B220")
  TCD = generate.UMAP(celldata = celldata, clusters = TCellClusters,
                      markers = TCellMarkers, nsample = 5000, filename = "TCD.csv")
  
  # Macrophages & Dendritic cells
  MacDendCellClusters = get.clusters(patterns = c("Macrophage", "Dendritic"))
  MacDendCellColumns = c("MI_MHCII", "MI_F480", "MI_CD103", "MI_PDL1", "MI_CD86",
                         "MI_CD68", "MI_CD206", "MI_pS6", "MI_CD11c", "MI_CD11b",'MI_Sirpa',"MI_CD44") #, "MI_LY6G",  "MI_CD44")
  MacDendCD = generate.UMAP(celldata = celldata, clusters = MacDendCellClusters,
                            markers = MacDendCellColumns, nsample = 15000,
                            filename = "MacDendCD.csv")
  
  # Fibroblasts & endothelial cells
  FibEndoClusters = get.clusters(patterns = c("Fibroblasts", "Endothelium"))
  FibEndoColumns = c("MI_aSMA", "MI_PDL1",  "MI_Vimentin",
                      "MI_PECAM")
  FibEndoCD = generate.UMAP(celldata = celldata, clusters = FibEndoClusters,
                            markers = FibEndoColumns, nsample = 10000,
                            filename = "FibEndoCD.csv")
}
# Newmyeloid clusters for umap
 MyeloidClusters = unique(myeloidCD$anncluster)
  MyeloidColumns = c("MI_MHCII", "MI_F480", "MI_CD103", "MI_PDL1", "MI_CD86",
                         "MI_CD68", "MI_CD206", "MI_pS6", "MI_CD11c", "MI_CD11b",'MI_Sirpa',"MI_CD44") #, "MI_LY6G",  "MI_CD44")
  Umap_myeloidCD = generate.UMAP(celldata = myeloidCD, clusters = MyeloidClusters,
                            markers = MyeloidColumns,
                            filename = "Umap_myeloidCD_newDC.csv")


```



##works,  has saved.  imegemetadata need to keep the same code with post preprocessing
```{r}
autosave.ROIs(filenames = unique(imagemetadata$filename),
              factor = "kinv25",imagemetadata = imagemetadata,
              directory = file.path(output_path, "Figures/Domain"),
              splitByROI = F)
```


##plot dots on all original image
```{r}
celldata<-read.csv(file=paste0(input_path,"2024_06_26_celldata.csv"))
ggplot(celldata, aes(x= Location_Center_X, y = -Location_Center_Y)) +
  geom_point(data=celldata,size = .1, color = "black") +
#geom_point(data = celldata[which(celldata$anncluster == "04_Dendritic cells"& (celldata$MI_MHCII>0.9|neighb1$MI_CD11c)),],
#              #      | celldata$MI_PECAM >=0.375)),],      
#           size = 0.005, colour = "yellow") +
  geom_point(data = celldata[which(celldata$annotation =="Dendritic cells"),], size = 0.01, colour = "red")+
   geom_point(data = celldata[which(celldata$annotation == "T cells CD8"),], size = 0.01, colour = "yellow")+
  theme_classic() +
  theme(axis.line=element_blank(),
        axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        legend.title = element_blank(),
        legend.text = element_text(size = 8),
        panel.background=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        strip.text = element_text(size = 8),
        strip.background = element_rect(),
        panel.spacing = unit(0, "mm"),
        plot.background = element_blank())+
  facet_wrap(~ROI_ID,nrow=3)
```
##plot dots on one specific original image
```{r}
ROI13<-subset(celldata, ROI_ID == "03_MRTX+PD1+CTLA-4") # "04_MRTX+PD1+CTLA-4"
ggplot(ROI13, aes(x= Location_Center_X, y = -Location_Center_Y)) +
  geom_point(size = 0.5, color = "black") +
  geom_point(data = ROI13[which(ROI13$annotation == "Dendritic cells"),], size = 0.5, colour = "red")+
   geom_point(data = ROI13[which(ROI13$annotation == "T cells CD8"),], size = 0.5, colour = "yellow")+
  theme_classic() +
  theme(axis.line=element_blank(),
        axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        legend.title = element_blank(),
        legend.text = element_text(size = 8),
        panel.background=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        strip.text = element_text(size = 8),
        strip.background = element_rect(),
        panel.spacing = unit(0, "mm"),
        plot.background = element_blank())+
  facet_wrap(~ROI_ID,nrow=1)


```








#plot annotated heatmap
```{r}
# make a new code to fit jan-piter's code about "kinv33"to anncluster
##get clusters from anncluster because of the reclustered myeloid
celldata <- celldata %>%
  dplyr::mutate(k30 = as.integer(gsub("(\\d{1,2}).*","\\1",anncluster)))
plot.heatmap2(celldata, lin_markers, "k30", title = "Cluster k30",
              Colv = T,Rowv = T, srtCol = 90, barplot = c("ExpGroup","ROI_ID"))



plot.heatmap2(celldata, lin_markers, "kinv33", title = "Cluster k30",
              Colv = T,Rowv = T, srtCol = 90, barplot = c("ExpGroup","ROI_ID"))

plot.heatmap2(celldata, lin_markers, "anncluster", title = "Annotated clusters",
              margins = c(10, 5), srtCol = 60, Colv = T,Rowv = T)
plot.heatmap2(celldata, all_markers, "anncluster", title = "Annotated clusters",
              margins = c(10, 5), srtCol = 60, Colv = F,Rowv = T)
plot.heatmap2(celldata, all_markers, "anncluster", title = "Annotated clusters",
              margins = c(10, 5), srtCol = 60, Colv = F,scale = "row")
plot.heatmap2(celldata, lin_markers, "annotation",
              title = "Cell type expression pattern",
              margins = c(9, 5), srtCol = 55, transpose = F,scale = "column")
plot.heatmap2(celldata, all_markers, "annotation",
              title = "Cell type expression pattern",
              margins = c(8, 9), srtCol = 55, transpose = T,scale = "column")
```





##Plot a heatmap to compare the annotated clustering with the original annotation:

```{r}
plot.heatmap2(celldata, lin_markers, "kinv33", title = "Cluster k30",Rowv = T,
              Colv = T, srtCol = 90, barplot = c("ExpGroup"))
plot.heatmap2(celldata, lin_markers, "kinv33", title = "Cluster k30",Rowv = T,
              Colv = T, srtCol = 90, barplot = c("ROI_ID"))



plot.heatmap2(celldata, lin_markers, "anncluster", title = "Annotated clusters",
              margins = c(10, 5), srtCol = 60, Colv = T,Rowv = T)
plot.heatmap2(celldata, all_markers, "anncluster", title = "Annotated clusters",
              margins = c(10, 5), srtCol = 60, Colv = T,Rowv = T)
plot.heatmap2(celldata, lin_markers, "annotation",
              title = "Cell type expression pattern",
              margins = c(10, 9), srtCol = 55, transpose = T,scale = "row",Colv = T,Rowv = T)
plot.heatmap2(celldata, all_markers, "annotation",
              title = "Cell type expression pattern",
              margins = c(10, 9), srtCol = 55, transpose = T,scale = "row",Colv = T,Rowv = T)
plot.heatmap2(celldata, lin_markers, "annotation",
              title = "Cell type expression pattern",
              margins = c(10, 9), srtCol = 55, transpose = T,Colv = T,Rowv = T)
plot.heatmap2(celldata, all_markers, "annotation",
              title = "Cell type expression pattern",
              margins = c(10, 9), srtCol = 55, transpose = T,Colv = T,Rowv = T)


##save as pdf 

output_dir<-"/mnt/Data1/groupfebe/runs/Xiaofei/Megan_IMC/Projectfile/Results/Figures_output/Heatmap"
pdf(file = paste0(output_dir,"/all_makers_expression.pdf"),width = 8,height = 6)
plot.heatmap2(celldata, all_markers, "annotation",
              title = "Cell type expression pattern",
              margins = c(10, 9), srtCol = 55, transpose = T,Colv = T,Rowv = T)
dev.off()
pdf(file = paste0(output_dir,"/lin_makers_expression.pdf"),width = 8,height = 6)
plot.heatmap2(celldata, lin_markers, "annotation",
              title = "Cell type expression pattern",
              margins = c(10, 9), srtCol = 55, transpose = T,Colv = T,Rowv = T)
dev.off()

```
## add  a new column cell_type for making heatmap or plot 
```{r}
celldata$cell_type<-celldata$annotation
celldata$cell_type[celldata$cell_type=="Macrophages F480"]<-"Macrophages"
celldata$cell_type[celldata$cell_type=="Macrophages CD68"]<-"Macrophages"
celldata$cell_type[celldata$cell_type=="Macrophages CD11b"]<-"Macrophages"
celldata$cell_type[celldata$cell_type=="cDC1 Dendritic cells"]<-"Dendritic cells"
celldata$cell_type[celldata$cell_type=="T cells CD4"]<-"CD4 T cells"
celldata$cell_type[celldata$cell_type=="T cells CD8"]<-"CD8 T cells"
celldata$cell_type[celldata$cell_type=="Tregs"]<-"Treg cells"
```



## CLuster k25  27 clusters distribution 
```{r tSNE for cluster distribution}
kinvlist = grep("kinv", names(celldata), value = T)
klist = paste0("k", round(1000/ as.numeric(str_remove(kinvlist, "kinv"))))
plot.tSNE(celldata, c("tSNE1", "tSNE2"), "kinv40", title = "Clusters k25",
          savetSNE = F)
```


## Visualize differences in clustering annotation with a t-SNE:
##Visulazation after annotation 
```{r}
plot.tSNE(celldata, c("tSNE1", "tSNE2"), "ExpGroup",
          title = "Experimental model", labelClusters = F)
plot.tSNE(celldata, c("tSNE1", "tSNE2"), "ROI_ID",
          title = "ROI ID", labelClusters = F)

plot.tSNE(celldata, c("tSNE1", "tSNE2"), "annotation", labelClusters = F,
          unique(celldata$annotation), title = "Annotated cell types",
          savetSNE = T)

plot.tSNE(celldata, c("tSNE1", "tSNE2"), "cell_type", labelClusters = F,
          unique(celldata$cell_type), title = "Annotated cell_types",
          savetSNE = T)
```

# febe plot dots on all original image or on specific image for checking the cutoff
##plot dots on all original image
```{r}
plot.scatter(celldata, "kinv33", 27, "MHCcII", "CD11b")
ggplot(celldata, aes(x= Location_Center_X, y = -Location_Center_Y)) +
  geom_point(data=celldata,size = .1, color = "black") +
geom_point(data = celldata[which(celldata$anncluster=="03B_Macrophages other"),],
           #           | celldata$MI_EpCAM >=0.375)),],      
           size = 0.01, colour = "yellow") +
  #geom_point(data = celldata[which(celldata$kinv33 =="11"& celldata$MI_CD44>0.4),], size = 0.01, colour = "red")+

  #geom_point(data = celldata[which(celldata$anncluster =="02A_Tumour" & celldata$MI_CK19 >0.2),], size = 0.01, colour = "red")+

  theme_classic() +
  theme(axis.line=element_blank(),
        axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        legend.title = element_blank(),
        legend.text = element_text(size = 8),
        panel.background=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        strip.text = element_text(size = 8),
        strip.background = element_rect(),
        panel.spacing = unit(0, "mm"),
        plot.background = element_blank())+
  facet_wrap(~ROI_ID,nrow=3)
```
##plot dots on one specific original image
```{r}
ROI13<-subset(celldata, ROI_ID == "13_Necroptosis") # 12_Vehicle 13_Necroptosis
ggplot(ROI13, aes(x= Location_Center_X, y = -Location_Center_Y)) +
  geom_point(size = 0.5, color = "black") +
  geom_point(data = ROI13[which(ROI13$MI_CD206>0.15),], size = 1, colour = "red")+
  theme_classic() +
  theme(axis.line=element_blank(),
        axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        legend.title = element_blank(),
        legend.text = element_text(size = 8),
        panel.background=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        strip.text = element_text(size = 8),
        strip.background = element_rect(),
        panel.spacing = unit(0, "mm"),
        plot.background = element_blank())+
  facet_wrap(~ROI_ID,nrow=1)


```



## plot scatter check, dont run here 
```{r}

```

## Barplot celltype percentage

Create a barplot displaying the percentage of each celltypes to the total
amount of cells in one experimental group.

```{r stacked barplot function}
# Plot cell type percentage ----------------------------------------------------
# Description : Creates a stacked barplot showing the distribution of cells
#               across the various levels in 'factor' for each experimental group.
# ---Parameters---
# celldata  --> [dataframe] a dataframe with cells annotated by a factor
# factor    --> [string] the name of a column of celldata that acts as a factor
# title     --> [string] the title of the stacked barplot
# colours   --> [vector] a named vector with colours, named with the levels of the factor 
# << Output >>  [image] The generated stacked barplot
plot.celltypeperc = function(celldata, factor, title, colours) {
  directory = create.resultsdir(path = output_path, name = "Stacked_barplots")
  filename = paste0(title, ".jpeg")
  factor = sym(factor)
  
  ggplot(celldata, aes(x = ExpGroup, fill = !!factor)) +
    geom_bar(position = "fill", colour = "black") +
    scale_fill_manual(values = colours) +
    scale_y_continuous(labels = scales::percent) + theme_classic() +
    scale_x_discrete(limits=c("MRTX+PD1","MRTX+PD1+CTLA-4","ETP")) +
    theme(axis.text.x = element_text(angle = 60, size =14, hjust = 1),
          axis.text.y = element_text(size =14),
          plot.title = element_text(size = 6),
          legend.title = element_blank(),
          legend.text = element_text(size = 14),
          axis.title = element_blank())
  
  ggsave(device = "jpeg", width = 2000, height = 2000, units = "px", dpi = 300,
         path = directory, filename = filename)
}

# Match clusters ---------------------------------------------------------------
# Description : Returns the levels in celldata$anncluster that match with one
#               of the strings in patterns
# ---Parameters---
# patterns  --> [vector] a vector with strings that match with the desired clusters
# << Output >>  [vector] a vector with the names of the matched clusters
get.clusters = function(patterns) {
  clusters = sapply(X = patterns, FUN = grep, x = unique(celldata$anncluster), value = T)
  clusters = unlist(clusters, use.names = F)
  
  return(clusters)
} 

# Get matched colours ----------------------------------------------------------
# Description : Returns a named vector with colours for each name in 'levels'
# ---Parameters---
# levels  --> [vector] a vector with names to which colours should be assigned
# << Output >>  [vector] a named vector of colours where names equals 'levels'
matchColours = function(levels) {
  colours = hue_pal()(length(levels))
  colours = setNames(colours, levels)
  return(colours)
}

plot.celltypeperc(celldata = celldata, title = "Stacked_metaclusters")
plot.celltypeperc(celldata = celldata[!celldata$annotation %in% c("Tumour", "Undefined"),],
                  title = "Stacked_without_tumour")
plot.celltypeperc(celldata = celldata[!celldata$annotation %in% c("Tumour", "Epithelium", "Fibroblasts", "Endothelium","Undefined"),],
                  title = "Stacked_immune_cells")
```

# UMAP

```{r plot UMAP function}
# Generate new UMAP ------------------------------------------------------------
# Description : Generate a new UMAP with the selected clusters and markers
# ---Parameters---
# celldata  --> [dataframe] a dataframe with mean intensity data by marker
# clusters  --> [vector] a vector with levels of the factor 'anncluster' that should be included
# markers   --> [vector] a vector with column names of celldata with marker
#               expression data that should be used to generate the UMAP
# nsample   --> [numeric] the number of cells that should be subsampled from celldata
# << Output >>  [dataframe] the selected subset of celldata with UMAP coordinates
generate.UMAP = function(datatable, clusters, markers, nsample, filename) {
  data = datatable[datatable$anncluster %in% clusters,]
  data = data[sample(x = nrow(data), size = nsample),]
  umapResults = umap(data[,markers], verbose = T, n_neighbors = 10)
  data$umap1 = umapResults$layout[,1]
  data$umap2 = umapResults$layout[,2]
  write.csv(x = data, file = file.path(output_path, filename), row.names = T)
  return(data)
}

# Plot UMAP -------------------------------------------------------------------
# Description : Plot cells on a plane according to their UMAP coordinates,
#               coloured according to levels of factor 
# ---Parameters---
# UMAPdf    --> [dataframe] a dataframe with UMAP data.
# factor    --> [string] name of a column of 'UMAPdf' that acts as a factor
# celltype  --> [string] a string that represents the type of cells in the UMAP
# lgndttl   --> [string] the title of plot legend
# << Output >>  [image] the generated UMAP
plot.UMAP = function(UMAPdf, factor, celltype, lgndttl) {
  factor = sym(factor)
  
  p = ggplot(UMAPdf, aes(x=umap1, y=umap2, colour = !!factor)) + 
    geom_point(size = 0.3) +
    guides(colour = guide_legend(override.aes = list(size=5), title = lgndttl)) +
    scale_color_manual(values = cluster_cols) + ggtitle(celltype) + 
    theme_classic() + labs(x = "UMAP 1", y = "UMAP 2") +
    theme(plot.background = element_blank(),
          axis.title = element_text(size = 15),
          legend.text = element_text(size = 15),
          legend.title = element_text(size = 17, face = "bold"),
          plot.title = element_text(hjust = 0.5, size = 22, face = "bold"),
          panel.grid = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank())
  
  celltype = str_replace_all(celltype, " ", "_")
  lgndttl = str_replace_all(lgndttl, " ", "_")
  filename = paste0("UMAP_", celltype, "_", lgndttl, ".jpeg")
  
  directory = create.resultsdir(path = output_path,
                                name = paste0(celltype, "_UMAP"))
  
  ggsave(filename = filename, path = directory,
         bg = "white", width = 2100, height = 1700, units = "px")
  
  print(p)
}

# Plot UMAP -------------------------------------------------------------------
# Description : Plot cells on a plane according to their UMAP coordinates,
#               coloured according to their expression of 'marker'  
# ---Parameters---
# UMAPdf    --> [dataframe] a dataframe with UMAP data.
# marker    --> [string] name of a column of 'UMAPdf' with expression data of a marker
# celltype  --> [string] a string that represents the type of cells in the UMAP
# << Output >>  [image] the generated UMAP
plot.UMAPmarker = function(UMAPdf, marker, celltype) {
  colour = sym(marker)
  title = paste(celltype, str_remove(marker, "MI_"), "expression")
  p = ggplot(UMAPdf, aes(x=umap1, y=umap2, colour = !!colour)) + 
    geom_point(size = 0.3) + ggtitle(title) + 
    theme_classic() + labs(x = "UMAP 1", y = "UMAP 2") +
    theme(plot.background = element_blank(),
          axis.title = element_text(size = 15),
          legend.text = element_text(size = 15),
          legend.title = element_blank(),
          plot.title = element_text(hjust = 0.5, size = 22, face = "bold"),
          panel.grid = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank()) +
    scale_color_distiller(palette = "Spectral", limits = c(0,3))
  
  celltype = str_replace_all(celltype, " ", "_")
  filename = paste0("UMAP_", celltype, "_", str_remove(marker, "MI_"), ".jpeg")
  
  directory = create.resultsdir(path = output_path,
                                name = paste0(celltype, "_UMAP"))
  
  ggsave(filename = filename, path = directory,
         bg = "white", width = 2100, height = 1700, units = "px")
  
  print(p)
}

cluster_cols = c("#00BA38", "#619CFF", "#BF2020", "#BFBF00","#E76BF3", "#00BFDF", "orange")
```

## plot marker expression (Tcolumns)  on T-umap 
```{r }
TCD = read.csv(paste0(input_path, "/TCD.csv"))
TColumns = c("MI_CD103", "MI_CD3e", "MI_CD45", "MI_CD4", "MI_CD8a", "MI_Foxp3",
                   "MI_Ki67", "MI_LAG3", "MI_PD1","MI_Casp3","MI_B220")
for (i in TColumns){
  marker = str_remove(i, "MI_")
  p = ggplot(TCD,
             aes(x=umap1, y=umap2, colour = get(i))) + 
    geom_point(size = 0.3) + ggtitle(paste0("T cell ", marker, " expression")) + 
    theme_classic() + labs(x = "UMAP 1", y = "UMAP 2") +
    theme(plot.background = element_blank(),
          axis.title = element_text(size = 15),
          legend.text = element_text(size = 15),
          legend.title = element_blank(),
          plot.title = element_text(hjust = 0.5, size = 25, face = "bold"),
          panel.grid = element_blank(),
          #axis.text = element_blank(),
          axis.ticks = element_blank()) +
    scale_color_distiller(palette = "Spectral", limits = c(0,1))
  ggsave(filename = paste0("UMAP_T_cell_", marker, "_expression.jpeg"),
         path = file.path(output_path, "TCellUMAP"), bg = "white",
         width = 2100, height = 1700, units = "px")
  print(p)
}

```

##plot marker expression (MACcolumns)  on MAC-umap 
```{r}
MacDendCellColumns = c("MI_MHCII", "MI_F480", "MI_CD103", "MI_PDL1", "MI_CD86",
                       "MI_CD68", "MI_CD206", "MI_pS6", "MI_CD11c", "MI_CD11b","MI_CD44")
for (i in MacDendCellColumns){
  marker = str_remove(i, "MI_")
  p = ggplot(MacDendCD,
             aes(x=umap1, y=umap2, colour = get(i))) + 
    geom_point(size = 0.3) + ggtitle(paste0("Myeloid ", marker, " expression")) + 
    theme_classic() + labs(x = "UMAP 1", y = "UMAP 2") +
    theme(plot.background = element_blank(),
          axis.title = element_text(size = 15),
          legend.text = element_text(size = 15),
          legend.title = element_blank(),
          plot.title = element_text(hjust = 0.5, size = 25, face = "bold"),
          panel.grid = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank()) +
    scale_color_distiller(palette = "Spectral", limits = c(0,3))
  ggsave(filename = paste0("UMAP_Myeloid_", marker, "_expression.jpeg"),
         path = file.path(output_path, "MyeloidUMAP"), bg = "white",
         width = 2100, height = 1700, units = "px")
  print(p)
}
ggplot(MacDendCD,aes(x=umap1, y=umap2, colour = anncluster))+geom_point()
ggplot(MacDendCD,aes(x=umap1, y=umap2, colour = ExpGroup))+geom_point()
```


## T cell

```{r}
TCellMarkers = c("MI_CD103", "MI_CD3e", "MI_CD45", "MI_CD4", "MI_CD8a", "MI_Foxp3",
                 "MI_Ki67", "MI_LAG3", "MI_PD1", "MI_Casp3")
ggplot(TCD,aes(x=umap1, y=umap2, colour = anncluster))+geom_point()
ggplot(TCD,aes(x=umap1, y=umap2, colour = annotation))+geom_point()
ggplot(TCD,aes(x=umap1, y=umap2, colour = ExpGroup))+geom_point()
plot.UMAP(UMAPdf = TCD, factor = "anncluster", celltype = "T cells", lgndttl = "Cluster")
plot.UMAP(UMAPdf = TCD, factor = "ExpGroup", celltype = "T cells", lgndttl = "Experimental group")
sapply(X = TCellMarkers, FUN = plot.UMAPmarker, UMAPdf = TCD, celltype = "T cells")
```

## Macrophage

```{r}
MacDendCellColumns = c("MI_MHCII", "MI_F480", "MI_CD103", "MI_PDL1", "MI_CD86",
                       "MI_CD68", "MI_CD206", "MI_pS6", "MI_CD11c", "MI_CD11b") #, "MI_LY6G",  "MI_CD44")
ggplot(Umap_myeloidCD_newDC,aes(x=umap1, y=umap2, colour = anncluster))+geom_point()
ggplot(Umap_myeloidCD_newDC,aes(x=umap1, y=umap2, colour = annotation))+geom_point()
ggplot(Umap_myeloidCD_newDC,aes(x=umap1, y=umap2, colour = ExpGroup))+geom_point()

plot.UMAP(UMAPdf = Umap_myeloidCD_newDC, factor = "anncluster",
          celltype = "Macrophages & Dendritic cells", lgndttl = "Cluster")
plot.UMAP(UMAPdf = Umap_myeloidCD_newDC, factor = "ExpGroup",
          celltype = "Macrophages & Dendritic cells", lgndttl = "Experimental group")
sapply(X = MacDendCellColumns, FUN = plot.UMAPmarker, UMAPdf = Umap_myeloidCD_newDC,
       celltype = "Macrophages & Dendritic cells")
```


# Draw ROIs

Plot each cell by x,y coordinate and color according to e.g. annotation
## metanr = grep(filenames[i], imagemetadata$filename) this code is wrong need to fix later
```{r draw-ROI functions, include=FALSE}
draw.ROI = function(filename, factor) {
  factor = sym(factor)
  celldatasubset = celldata[celldata$filename == filename,]
  p = ggplot(celldatasubset, aes(x= Location_Center_X, y = -Location_Center_Y,
                                 color = factor(!!factor))) +
    geom_point(size = .2) +
    theme_classic() +
    theme(axis.line=element_blank(),
          axis.text.x=element_blank(),
          axis.text.y=element_blank(),
          axis.ticks=element_blank(),
          axis.title.x=element_blank(),
          axis.title.y=element_blank(),
          legend.title = element_blank(),
          legend.text = element_text(size = 8),
          panel.background=element_blank(),
          panel.grid.major=element_blank(),
          panel.grid.minor=element_blank(),
          strip.text = element_text(size = 8),
          strip.background = element_rect(),
          panel.spacing = unit(0, "mm"),
          plot.background = element_blank()) +
    guides(colour = guide_legend(override.aes = list(size=5)))
  
  return(p)
}

autosave.ROIs = function(filenames, factor, directory, imagemetadata, splitByROI = T) {
  date = str_sub(gsub("-", "", Sys.Date()), -6, -1)
  for (i in 1:length(filenames)) {
    
    if(splitByROI){
      file_directory = file.path(directory, filenames[i])
      if (!dir.exists(file_directory)) {
        dir.create(file_directory)
      }
      
      file_name = paste0(date, "_", factor, ".jpeg")
    } else {
      file_name = paste0(date, "_", filenames[i], ".jpeg")
      file_directory = directory
    }
    
    metanr = grep(filenames[i], imagemetadata$filename)
    p = draw.ROI(filenames[i], factor = factor)
    
    ggsave(plot = p, device = "jpeg", bg = "white",
           width = imagemetadata$width[metanr],
           height = imagemetadata$height[metanr],
           dpi = 300, units = "px", path = file_directory, filename = file_name)
  }
}
#   copy from figures.rmd
# Plot ROI ---------------------------------------------------------------------
# Description : Creates an ROI plot by plotting the cells of the ROI on an x-y
#               plane according to their coordinates, and colours the cells
#               according to the levels of 'factor'.
# ---Parameters---
# celldata    --> [dataframe] a dataframe with cells annotated by a factor.
# ROIdata     --> [dataframe] a dataframe with the filename of each ROI and its
#                 height and width in pixels
# filename    --> [character] a level of the factor 'filename' in celldata
# factor      --> [string] the name of a column of celldata that acts as a factor
# splitByROI  --> [boolean] if T the ROI plot is saved in a subfolders named after the filename
# << Output >>    [image] The generated ROI plot
plot.ROI = function(celldata, ROIdata, filename, factor, splitByROI = F) {
  factor = sym(factor)
  celldatasubset = celldata[celldata$filename == filename,]
  p = ggplot(celldatasubset, aes(x = Location_Center_X, y = -Location_Center_Y,
                                 color = factor(!!factor))) +
    geom_point(size = .2) + theme_classic() +
    theme(axis.line = element_blank(), axis.text.x=element_blank(),
          axis.text.y = element_blank(), axis.ticks=element_blank(),
          axis.title.x = element_blank(), axis.title.y=element_blank(),
          legend.title = element_blank(), legend.text = element_text(size = 8),
          panel.background = element_blank(), panel.grid.major=element_blank(),
          panel.grid.minor = element_blank(), strip.text = element_text(size = 8),
          strip.background = element_rect(), panel.spacing = unit(0, "mm"),
          plot.background = element_blank()) +
    guides(colour = guide_legend(override.aes = list(size=5)))
  
  print(p)
  
  metanr = grep(filename, ROIdata$filename)
  directory = create.resultsdir(path = output_path, name = "Domain")
  name = paste0(filename, ".jpeg")
  
  if(splitByROI){
    directory = file.path(directory, filename)
    name = paste0(factor, "_ROI.jpeg")
    if (!dir.exists(directory)) dir.create(directory)
  }
  
  ggsave(plot = p, device = "jpeg", bg = "white", dpi = 300, units = "px",
         width = ROIdata$width[metanr], height = ROIdata$height[metanr],
         path = directory, filename = name)
}

```
##save ROIs Doesnt work for now,  check later
```{r}
autosave.ROIs(filenames = unique(imagemetadata$filename),
              factor = "annotation",imagemetadata = imagemetadata,
              directory = file.path(output_path, "Figures/Domain"),
              splitByROI = F)
```


# Stacked Barplot (celltype percentage--Important)
##Create a barplot displaying the percentage of each celltypes to the total
amount of cells in one experimental group.
```{r}
plot.celltypeperc = function(celldata, factor, title, colours) {
  directory = create.resultsdir(path = output_path, name = "Stacked_barplots")
  filename = paste0(title, ".jpeg")
  factor = sym(factor)
  
  ggplot(celldata, aes(x = ExpGroup, fill = !!factor)) +
    geom_bar(position = "fill", colour = "black") +
    scale_fill_manual(values = colours) +
    scale_y_continuous(labels = scales::percent) + theme_classic() +
    scale_x_discrete(limits=c("MRTX+PD1","RTX+PD1+CTLA-4")) +
    theme(axis.text.x = element_text(angle = 60, size =14, hjust = 1),
          axis.text.y = element_text(size =14),
          plot.title = element_text(size = 6),
          legend.title = element_blank(),
          legend.text = element_text(size = 14),
          axis.title = element_blank())
  
  ggsave(device = "jpeg", width = 2000, height = 2000, units = "px", dpi = 300,
         path = directory, filename = filename)
}

# Match clusters ---------------------------------------------------------------
# Description : Returns the levels in celldata$anncluster that match with one
#               of the strings in patterns
# ---Parameters---
# patterns  --> [vector] a vector with strings that match with the desired clusters
# << Output >>  [vector] a vector with the names of the matched clusters
get.clusters = function(patterns) {
  clusters = sapply(X = patterns, FUN = grep, x = unique(celldata$anncluster), value = T)
  clusters = unlist(clusters, use.names = F)
  
  return(clusters)
}

# Get matched colours ----------------------------------------------------------
# Description : Returns a named vector with colours for each name in 'levels'
# ---Parameters---
# levels  --> [vector] a vector with names to which colours should be assigned
# << Output >>  [vector] a named vector of colours where names equals 'levels'
matchColours = function(levels) {
  colours = hue_pal()(length(levels))
  colours = setNames(colours, levels)
  return(colours)
}
plot.celltypeperc(celldata = celldata, title = "Stacked_metaclusters")
plot.celltypeperc(celldata = MyeCD, title = "Stacked_macclusters")
plot.celltypeperc(celldata = celldata[!celldata$annotation %in% c("Tumour", "Undefined"),],
                  title = "Stacked_without_tumour")
plot.celltypeperc(celldata = celldata[!celldata$annotation %in% c("Tumour", "Epithelium", "Fibroblasts", "Endothelium","Undefined"),],
                  title = "Stacked_immune_cells")

```


date <- Sys.Date()
output_file = paste0(output_path, date, "_celldata.csv")
output_file
write.csv(celldata, file = output_file, row.names = F) 
